---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Ye Dam Yi"
date: "03/02/2021"
output: github_document
---

### Load packages and data

```{r install-packages, message = FALSE}
# install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")
```

```{r load-packages}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
str(states)
```

### Exercise 1

There are 6 columns and 1,643 rows in dennys. The columns represent location information about Denny's. The rows represent different Dennis' branches. 

```{r dennys-dimensions}
glimpse(dennys)
nrow(dennys)
ncol(dennys)
```


### Exercise 2

There are 6 columns and 909 rows in laquinta. The columns represent location information about La Quinta branches. The rows represent each La Quinta branch. 

```{r laquinta-dimensions}
glimpse(laquinta)
nrow(laquinta)
ncol(laquinta)
```


### Exercise 3

There are La Quinta's locations outside of the US. The countries include Chile, Colombia, Honduras, Canada, and Mexico. 

There are no Denny's locations outside of the US. 

### Exercise 4

Ways to determine whether or not any establishment has any locations outside the U.S. 

1. Determine whether the state is one of the states for the US. 
2. Determine whether the longitude and the latitude are within the ranges of those for the US. 

### Exercise 5

```{r rename-datasets}
dn <- dennys
lq <- laquinta

```

Filter for states that are not in states$abbreviation.
```{r filter-not-states}
dn %>% 
  filter(!(state %in% states$abbreviation))
```

### Exercise 6

Add a country variable to the Denny's dataset and set all observations equal to "United States."

```{r set-country}
dn %>% 
  mutate(country = "United States")
```

### Exercise 7

Find the La Quinta locations that are outside the US, and figure out which country they are in.

```{r lq-outside-us}
lq_outside_us <- lq %>% 
  filter(!(state %in% states$abbreviation))
lq_outside_us

lq_outside_us <- lq_outside_us %>% 
  mutate(country = c("Mexico", "Mexico", "Mexico", "Mexico", "Colombia", "Mexico", "Mexico", "Canada", "Mexico", "Mexico", "Mexico", "Mexico", "Honduras", "Canada"))
lq_outside_us

```
### Exercise 8
```{r country-variable-lq}
lq <- lq %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States", # state.abb is a dataset contained in R
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"
    )
  )
glimpse(lq)

lq <- lq %>%
  filter(country == "United States")
```

### Exercise 9
```{r count-by-state}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>% 
  arrange(n)

lq %>% 
  count(state) %>% 
  inner_join(states, by = c("state" = "abbreviation"))%>% 
  arrange(n)

```
California has the most Denny's locations, and Delaware has the fewest. It's surprising given that Denny's serves traditional American food, such as burgers and breakfast items, which I would think are popular in the South than in the Western coast.  

Texas has the most La Quinta's locations, and Maine has the fewest. It's not that surprising because La Quinta's a hotel, and it's likely to be in a place where people would be traveling a lot. The fact that there are the most La Quinta's locations in Texas, Florida, California, and Georgia seems to be in line with that.It's also not suprising because California has the most Denny's locations. 

### Exercise 10
```{r per-area}
# we want to put together the two datasets into one data frame
dn <- dn %>% 
  mutate(establishment = "Denny's")
lq <- lq %>% 
  mutate(establishment = "La Quinta")

# use bind_rows to bind the datasets because they have the same columns
dn_lq <- bind_rows(dn, lq)

# plot the location of the two establishments

cb.palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point() +
  scale_color_manual(values = cb.palette)

```


### Exercise 11
``` {r dn_lq-nc}
dn_lq %>% 
  filter(state == "NC") %>% 
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = .66) +
  scale_color_manual(values = cb.palette)

```

Visually, Mitch Hedberg's joke partially holds in North Carolina. Some of La Quinta's locations in the state are close to Denny's while others are not. 

### Exercise 12
```{r dn_lq-tx}
dn_lq %>% 
  filter(state == "TX") %>% 
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = .66) +
  scale_color_manual(values = cb.palette)

```

Mitch Hedberg’s joke appears to hold in Texas. Most of Denny's and La Quinta's locations in Texas are next to each other. 


