---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Ye Dam Yi"
date: "02/24/2021"
output: github_document
---

### Load packages and data

```{r install-packages, message = FALSE}
# install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")
```

```{r load-packages}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
str(states)
```

### Exercise 1

There are 6 columns and 1,643 rows in dennys. The columns represent location information about Denny's. The rows represent different Dennis' branches. 

```{r dennys-dimensions}
glimpse(dennys)
nrow(dennys)
ncol(dennys)
```


### Exercise 2

There are 6 columns and 909 rows in laquinta. The columns represent location information about La Quinta branches. The rows represent each La Quinta branch. 

```{r laquinta-dimensions}
glimpse(laquinta)
nrow(laquinta)
ncol(laquinta)
```


### Exercise 3

There are La Quinta's locations outside of the US. The countries include Chile, Colombia, Honduras, Canada, and Mexico. 

There are no Denny's locations outside of the US. 

### Exercise 4

Ways to determine whether or not any establishment has any locations outside the U.S. 

1. Determine whether the state is one of the states for the US. 
2. Determine whether the longitude and the latitude are within the ranges of those for the US. 

### Exercise 5

```{r rename-datasets}
dn <- dennys
lq <- laquinta

```

Filter for states that are not in states$abbreviation.
```{r filter-not-states}
dn %>% 
  filter(!(state %in% states$abbreviation))
```

### Exercise 6

Add a country variable to the Denny's dataset and set all observations equal to "United States."

```{r set-country}
dn %>% 
  mutate(country = "United States")
```

Add exercise headings as needed.

