---
title: "Lab 05 - Wrangling spatial data"
author: "Ye Dam Yi"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}

devtools::install_github("rstudio-education/dsbox")

```

```{r library}
library(tidyverse) 
library(dsbox) 
```
### Data declaration

```{r data}
dn <- dennys
lq <- laquinta
```

### Ex 1. Filter the Denny’s dataframe for Alaska (AK) and save the result as dn_ak. How many Denny’s locations are there in Alaska?

```{r ex1-dn-alaska}
dn_ak <- dn %>%
  filter(state == "AK") # filter for Denny's locations in AK
nrow(dn_ak)             # count the number of observations

```
There are three Denny's locations in Alaska. 

### Ex 2. Filter the La Quinta dataframe for Alaska (AK) and save the result as lq_ak. How many La Quinta locations are there in Alaska?
```{r ex2-lq-alaska}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are two La Quinta's locations in Alaska.


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
